---
title: "System 1 and System 2 Emotion 2021"
author: "Mikkel Wallentin"
date: "25 October 2021"
output:
  html_document: default
  pdf_document: default
---

##Emotion data
```{r}
#Loading libraries
library(pacman)
p_load(modelr,lmerTest,ggplot2,dplyr,stringr,tidyverse,plyr,pastecs,data.table,knitr,reshape)

#Loading all the datafiles for a defined path
#fill in path = "your data's directory"

temp = list.files(path = "emotion_systems_data", pattern = "*.csv", full.names = T)

#Applying read.delim to all files in temp, so all files is read into R
myfiles = lapply(temp,read.delim,sep=",",header = T)

#Binding list of files to a df
EMO_df=rbindlist(myfiles)
```


####Removing data without responses
```{r}
#Remove NAs
EMO_df=subset(EMO_df,EMO_df$rt!="NA")

#Keeping only RT's above 0 (in case something went wrong with off-sets)
EMO_df=subset(EMO_df,EMO_df$rt>0)
```

####Assessing the responses
```{r}
#Making a table showing how "emotional" each category of stimulus where judged to be (based on mean of responses, 0 = no emotion, 1 = emotion --> ideally TP1 and TP2 should be very close to 1)

#Can be used to assess whether the stimuli in the "emotional" categories (TP1 and TP2) are actually judged by the participants to contain emotion. 
aggdata_1 <-aggregate(EMO_df$emotion_judge,
    by=list(EMO_df$categories),
                    FUN=mean, na.rm=TRUE)

names(aggdata_1)<-c('Image category','% emotional')
aggdata_1

#Making a table with mean RT for each image category - eyeballing the result.
aggdata_2 <-aggregate(EMO_df$rt,
    by=list(EMO_df$categories),
                    FUN=mean, na.rm=TRUE)

names(aggdata_2)<-c('Image category','mean RT')
aggdata_2
```
##Please comment on this
System one is faster



##Is the difference in RT between categories significant?
```{r}
library (lmerTest)

#Turn the ID variable into a factor to check if there is more than one, i.e. if the script has loaded data from more than one participant/session
EMO_df$ID<-as.factor(EMO_df$ID)

#if the model is run on several participants, we choose a mixed-effects model (lmer)
if (length(unique(EMO_df$ID))>1) {
  emo_model=lmer(rt~categories + (1|ID),data=EMO_df)
}
# Else we choose an ordinary linear regression model (lm)
if (length(unique(EMO_df$ID))==1) {
emo_model=lm(rt~categories,data=EMO_df)
}

summary(emo_model)
```
##Comment on this
Both changes in reaction time is signifacantly different from zero



##Plot histogram of data (Checking some assumptions)
```{r}
#histogram
hist(EMO_df$rt,breaks=10)

#Q-Q-plot
qqnorm(EMO_df$rt)
```
##Comment on this



##Try with a log-transform, if the data is not normally distrubuted
```{r}
#make a log-transformation
EMO_df$logrt=log(EMO_df$rt)
#histogram
hist(EMO_df$logrt,breaks=10)
#Q-Q-plot
qqnorm(EMO_df$logrt)
```
##Comment on this
We can transform the data, to make analysis as if its normally distributed


##A side note on assumptions:

One can argue whether this variable (logrt) should be used instead of the regular rt. How much do you know about assumptions? 

In addition: Assumption of independence is also violated (because the experiment uses repeated measures) - we can take this into account using a mixed effects model (lmer) instead of a linear model (lm), as you may have learned/will learn in methods. 

```{r}
emo_model2 = lmerTest::lmer(logrt~categories+(1|ID),EMO_df)
summary(emo_model2)
#note: This will only run if you have data from multiple participants
```
##Comment on this
Very significant


##Plotting of data
```{r}
ggplot(EMO_df, aes(x = categories , y = rt)) +
  geom_boxplot() + labs(x = "Image Category", y = "Response time")


b=ggplot(EMO_df,aes(categories,rt))
b+geom_bar(stat = "summary",fun.y=mean, fill="gold1")+
  labs(x= "Image Category", y="Response Time",title="Response Time for Emotional Stimuli")
```
##Please write a paragraph summarising the following:

- The experiment 
- The results

##Prepare for class discussion. Discuss the following in your group and take notes:

- What might the hypothesis of this experiment be? 
- Can you find theoretical motivation for this hypothesis?
- What can or can’t we tell from the results?
- Could this experiment be modified to investigate another hypothesis?
- What can we use from this experiment? (What is clever in the experimental design?)
- How can we improve this experiment? (What is not so clever in the experimental design?)




