---
title: "Maximizing analysis 2022"
author: "Mikkel Wallentin"
date: "4 February 2022"
output:
  pdf_document: default
  html_document: default
---

## Load data
```{r}

pacman::p_load(readxl)
Max <- read.csv("Maximizing Regret Happiness Questionnaires (Responses) - Form Responses 1.csv")

#Rename variables for convenience
colnames(Max)<-c('Timestamp','Year','Gender','Max01','Max02','Max03','Max04','Max05','Max06','Max07','Max08','Max09','Max10','Max11','Max12','Max13','Regret01','Regret02','Regret03','Regret04','Regret05','Happ01','Happ02','Happ03','Happ04')
```

#### Notation. 
When indexing matrices or data frames, we can use square brackets like this [row,column], e.g. [1,2] means the first row of the second column. Leaving a place empty means "all", e.g. [,16] means all rows in column 16.

## Preproces data
```{r}
#Find maximizing quotient for each participant
Max$MaxqSum<-rowSums(Max[,4:16])
#The first variable of the Regret Questionnaire is phrased opposite of the others (value is to be subbracted from 8)

Max[,17]<-8-Max[,17]
Max$RegretSum<-rowSums(Max[,17:21])

#The last variable of the happiness questionnaire is phrased opposite of the others (value is to be subbracted from 8)
Max[,25]<-8-Max[,25]
Max$HappinessSum<-rowSums(Max[,22:25])
```


### Plot Maximizing sum against Regret sum

```{r}
pacman::p_load(ggplot2)
splot<-ggplot(Max, aes(MaxqSum,RegretSum))
#Plot scatterplot and a linear regression line
splot + geom_smooth(method='lm')+ geom_point(size=2) + labs(y="Regret Scale Score", x='Maximizing Scale Score')

```

### Plot Maximizing sum against Happiness sum

```{r}
library(ggplot2)
splot<-ggplot(Max, aes(MaxqSum,HappinessSum))
#Plot scatterplot and a linear regression line
splot + geom_smooth(method='lm')+ geom_point(size=2) + labs(y="Happiness Scale Score", x='Maximizing Scale Score')
#clean up
rm(splot)
```


##Statistical analysis

####Pearson correlation between Maximizing, Regret and Happiness scores
```{R}
#Pairwise correlation test between Maximization and Regret
Maxcor<-cor.test(Max$MaxqSum,Max$RegretSum)
Maxcor
#Pairwise correlation test between Maximization and Happiness
Maxcor<-cor.test(Max$MaxqSum,Max$HappinessSum)
Maxcor

#Pairwise correlation test between Regret and Happiness
Maxcor<-cor.test(Max$RegretSum,Max$HappinessSum)
Maxcor

#Clean up
rm(Maxcor)
```

#### Let's make a correlogram with summary scores
```{R}

# Use GGally to plot correlation matrix of summary scores
pacman::p_load(GGally)

ggpairs(Max[c(26:28)], title="Correlogram: Summary scores") 


# ggpairs shows us 
# a. The distribution of variables on the diagonal
# b. The correlation coefficients on the right
# c. scatter plot of variables on the left

```


#### Let's make a correlogram with all possible correlations
```{R}
#Change gender variable into numeric in order to use it in correlation
Max$Gender<-as.factor(Max$Gender)
Max$GenderN<-as.numeric(Max$Gender)

# Use GGally to plot correlation matrix of both individual questions and summaries
pacman::p_load(ggcorrplot)
# ggcorrplot colors the squares according to whether the correlation is positive or negative
ggcorrplot(cor(Max[c(4:29)]))
```

### Include significance test
```{R}
# Using p.mat, we can cross out the squares that are not significant at a given cutoff.
#cor_pmat(): Returns a matrix containing the p-values of correlations
P<-cor_pmat(Max[c(4:29)])
ggcorrplot(cor(Max[c(4:29)]), p.mat = P, sig.level=0.05)
```


#### Remake the correlogram with binary gender
```{R}
library(tidyverse)
#Filter variable to only contain males and females
Max2 <- Max %>% dplyr::filter(.,Gender == "Male" | Gender == "Female")

# Use GGally to plot correlation matrix of summary scores
library(GGally)
ggpairs(Max2[c(26:29)], title="Correlogram: Summary scores") 


# Use GGally to plot correlation matrix of both individual questions and summaries
library(ggcorrplot)
ggcorrplot(cor(Max2[c(4:29)]), p.mat = cor_pmat(Max2[c(4:29)]), sig.level=0.05)
```

# Code for voluntary extra questions

## A bit of power analysis for correlations

```{r}

#Statistical power is the probability of a hypothesis test of finding an effect if there is an effect to be found.

# Correlation can be viewed as an effect size measure.

# A correlation coefficient of 
#.20  RMPE recommended minimum effect size representing a ???practically??? significant effect for social science data (but smaller may in certain cases be interesting); 
#.50 a moderate correlation; 
#.80 is considered to represent a strong or large correlation 
#(https://psychology.okstate.edu/faculty/jgrice/psyc3214/Ferguson_EffectSizes_2009.pdf).
# 
# We can obtain sample size for a significant correlation at a given alpha level or the power for a given sample size using the function wp.correlation() from the R package webpower.

#A pairwise correlation test between Maximization and Happiness to exemplify
Maxcor<-cor.test(Max$Maxq,Max$Happiness)
Maxcor

#Using WebPower to find the power for detecting an effect with the current effect size and number of participants
library(WebPower)
wp.correlation(power=0.9, r= Maxcor$estimate)

#Make power estimates for different sample sizes
powercurve=wp.correlation(n=seq(4,10000,1), r=Maxcor$estimate, alternative = "two.sided") 
#Plot the power curve
plot(powercurve,type='l')

```

## Plotting Power curves for correlation coefficients and sample size

```{r}
#Code from here:
#https://www.statmethods.net/stats/power.html

# Plot sample size curves for detecting correlations of
# various sizes.

library(pwr)

# range of correlations
r <- seq(.1,.5,.01)
nr <- length(r)

# power values
p <- seq(.4,.9,.1)
np <- length(p)

# obtain sample sizes
samsize <- array(numeric(nr*np), dim=c(nr,np))
for (i in 1:np){
  for (j in 1:nr){
    result <- pwr.r.test(n = NULL, r = r[j],
    sig.level = .05, power = p[i],
    alternative = "two.sided")
    samsize[j,i] <- ceiling(result$n)
  }
}

# set up graph
xrange <- range(r)
yrange <- round(range(samsize))
colors <- rainbow(length(p))
plot(xrange, yrange, type="n",
  xlab="Correlation Coefficient (r)",
  ylab="Sample Size (n)" )

# add power curves
for (i in 1:np){
  lines(r, samsize[,i], type="l", lwd=2, col=colors[i])
}

# add annotation (grid lines, title, legend) 
abline(v=0, h=seq(0,yrange[2],50), lty=2, col="grey89")
abline(h=0, v=seq(xrange[1],xrange[2],.02), lty=2,
   col="grey89")
title("Sample Size Estimation for Correlation Studies\n
  Sig=0.05 (Two-tailed)")
legend("topright", title="Power", as.character(p),
   fill=colors)
```


## Factor analysis

#### Factor analysis takes data and uses internal correlations among variables to determine the inherent factors a multivariate dataset contains. Variables that are highly correlated belongs to the same factor.
```{r}


###  Prepare factor analysis

#Factor analysis on questions
library(psych)
#Use "parallel" function to find a suitable number of factors, based on the assumption that the variance explained in real factors should be higher than factors obtained using a random shuffle of the data.
parallel <- fa.parallel(Max[4:25], fm = 'minres', fa = 'fa',sim=FALSE)


### Make factor analysis

#Parallel function suggests nfact factors (see parallel$nfact), we will use minres and varimax rotation
MaxFa <- fa(Max[4:25],nfactors = parallel$nfact,rotate = "varimax",fm="minres")
print(MaxFa, cut=0.1, order=TRUE)

```