---
title: "FaceStroop Experiment Analysis: Part 1"
author: "Mikkel Wallentin"
date: "10/02/2022"
output: html_document
---
#Experiment

This is a donder-type reaction type experiment, meassuring our RT at different points of our circadian rythms, and when fasting.
The experiment also utilizes Stroop type stimuli, congurent/incongurent

##Get and clean FaceStroopExp data
```{r}
library(tidyverse)

#Get FaceStroopData
datadir<-"./FaceStroopExp_data"


#Find files
files<-list.files(datadir,pattern='^FaceStroop_.+?csv',full.names=TRUE)

#Prepare an empty data frame for the data (also removes old version)
dataFS<-data.frame()

#How many datasets were there
n_datasets_raw<-length(files)
#Prepare a variable to monitor how many datasets we keep
n_datasets<-0
#Prepare a variable to monitor how many points we originally had
n_datapoints_raw<-0

#Loop to go through all files in the list
for(iii in 1:n_datasets_raw){
  
  #remove old loaded file to not risk importing it multiple times
  if(exists('data_temp')) rm(data_temp)
  
  #Load data
  data_temp<-read.csv(files[iii])
  
  #Take trials that have an RT (remove intertrial fixations)
  data_temp<- subset(data_temp,rt!='null' & trial_type=='image-keyboard-response')
  
  #Make a variable with the total number of loaded data points, to know how many have been discarded 
  n_datapoints_raw<-n_datapoints_raw+length(data_temp[,1])
  
  #Select data where at least 20 trials have been completed
  if(length(data_temp[,1])>20){
  
  #Some early data did not have a hunger measurement. Add a hunger variable for them and make them NAs
  if("hunger" %in% names(data_temp)){} else data_temp$hunger=NA
  
  #Some very ugly code because of script bug (you are able to select more than one value for hours since sleep and eat). 
  #We split the string, convert it from a list and turn the first value into an integer (a number)
  if(is.integer(data_temp$H_since_Eat)==FALSE){
    data_temp$H_since_Eat<-as.integer(unlist(strsplit(data_temp$H_since_Eat[1],","))[1])
    }
  if(is.integer(data_temp$H_since_Sleep)==FALSE){
    data_temp$H_since_Sleep<-as.integer(unlist(strsplit(data_temp$H_since_Sleep[1],","))[1])
    }
  
  #Make sure that data have the same number of columns
  if(length(colnames(data_temp))==33){
    #Only include dataset with more than 70 percent correct responses
    if(sum(data_temp$correct=='true')/length(data_temp$correct)>0.70){
      data_temp$correct_p<-mean(as.numeric(data_temp$correct=='true'))
      
      #Turn RT into numeric variable
      data_temp$rt<-as.numeric(data_temp$rt)
      
      #Determine a cutoff for when response times are suspiciously long (e.g. 3 SDs above the mean)
      cutoff<-mean(data_temp$rt)+3*sd(data_temp$rt)
      #Filter data with too long RTs
      data_temp<- subset(data_temp,rt<cutoff)

      #If this is the first time around the loop, make the currently loaded data into the actual data
      if(length(colnames(dataFS))==0){
        dataFS=data_temp
        rm(data_temp)
        #counter to monitor included datasets
        n_datasets<-n_datasets+1
      }
      #Bind loaded data with actual data
      else {dataFS<-rbind(dataFS,data_temp)
      rm(data_temp)
        #counter to monitor included datasets
        n_datasets<-n_datasets+1
      }
    }
  }
  }
}





#A variable to monitor how many points we keep
n_datapoints<-length(dataFS[,1])
```


## Prepare variables for analysis
```{r}
#Make a variable that says if the condition was a Stroop condition or not (incongruency between word and face)
dataFS$stroop<-as.factor(dataFS$face!=dataFS$word)

#Limit the data to only correct answers. Response time for incorrect responses can be difficult to interpret.
#Note that this will make the data unsuitable for analyses on accuracy.
dataFS<- subset(dataFS,correct=='true')

#Turn id into factor - this can be used to monitor number of unique participants and to model random effects
dataFS$id <- dataFS$id %>% as.factor()

#Make a variable which has exp onset in hour and minutes of the day as decimal variable
dataFS$hour2<-dataFS$hour+(dataFS$minute)/60
```


## Plots


#### Let's make a correlogram with all numeric variables
```{R}
library(tidyverse)
x <- dataFS %>% dplyr::select(where(is.numeric))
# Use GGally to plot correlation matrix of both individual questions and summaries
library(ggcorrplot)
# ggcorrplot colors the squares according to whether the correlation is positive or negative
ggcorrplot(cor(x))

rm(x)
```

## Plotting response time as a function of hour
```{r}
library(ggplot2)


ggplot(dataFS,aes(x=hour2,y=rt))+geom_smooth()+geom_point()+labs(x='Time of day (hours)', y='Response time')

ggplot(dataFS,aes(x=hour2,y=correct_p))+geom_smooth()+geom_point()+labs(x='Time of day (hours)', y='Accuracy (percent)')

ggplot(dataFS,aes(x=hour2,y=mood))+geom_smooth(na.rm=TRUE)+geom_point()+labs(x='Time of day (hours)', y='Negative/Positive Mood')

ggplot(dataFS,aes(x=hour2,y=hunger))+geom_smooth(na.rm=TRUE)+geom_point()+labs(x='Time of day (hours)', y='HungrySatiated')
ggplot(dataFS,aes(x=hour2,y=fresh))+geom_smooth(na.rm=TRUE)+geom_point()+labs(x='Time of day (hours)', y='Tired/Fresh')

ggplot(dataFS,aes(x=H_since_Eat,y=hunger))+geom_smooth(na.rm=TRUE)+geom_point()+labs(x='Time since eating (hours)', y='Hungry/Satiated')

ggplot(dataFS,aes(x=H_since_Eat,y=mood))+geom_smooth(na.rm=TRUE)+geom_point()+labs(x='Time since eating (hours)', y='Negative/Positive Mood')


ggplot(dataFS,aes(x=hour2,y=mood,col=id))+geom_line() + stat_summary(aes(group = 1), geom = "point", fun.y = mean,shape = 17, size = 3)+labs(x='Time of day (hours)', y='Negative/Positive Mood')
```

## Example mixed effects models

### Response time

#### A simple model with only experimental variables

```{r}
library(lmerTest)

## Simple 2x2 model with stroop incongruency, face type and interaction and participants as random effects
modelFSrt1<-lmer(rt~stroop*face+(1|id),data=dataFS)
summary(modelFSrt1)
```

#### An inclusive model with all sorts of variables
```{r}
#Inclusive model
modelFSrt2<-lmer(rt~+stroop+face+hour2+trial_index+H_since_Sleep+H_since_Eat+mood+Gender+(1|id),data=dataFS)
summary(modelFSrt2)
```

### Mood

#### A model of mood given sleep and food
```{r}
#modelFSmood<-lmer(mood~H_since_Sleep*H_since_Eat+(1|id),data=dataFS)
#summary(modelFSmood)

#not repated mood measurements
```

### Wakefulness

#### A model of wakefulness given sleep and food
```{r}
modelFSfresh<-lmer(fresh~H_since_Sleep*H_since_Eat+(1|id),data=dataFS)
summary(modelFSfresh)
```

```{r}
sapply(dataFS,class)



```


```{r}






```




```{r}


scatterplot<- ggplot(dataFS,aes(mood,fresh))+
  geom_point(aes())+
  geom_smooth(method=lm,linetype="dashed",color="darkred")+
  theme(legend.position=c(0,1), legend.justification=c(0,1))


scatter_gender_reg<- ggplot(dataFS,aes(mood,fresh))+
  geom_point()+
  geom_smooth(method=lm,linetype="dashed")+
  theme(legend.position=c(0,1), legend.justification=c(0,1))


xdensity <- ggplot(dataFS, aes(mood)) + 
  geom_density(alpha=.5) + 
  theme(legend.position = "none")

ydensity <- ggplot(dataFS, aes(fresh)) + 
  geom_density(alpha=.5) + 
  theme(legend.position = "none")

blankPlot <- ggplot()+geom_blank(aes(1,1))+
  theme(plot.background = element_blank(), 
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(), 
   panel.border = element_blank(),
   panel.background = element_blank(),
   axis.title.x = element_blank(),
   axis.title.y = element_blank(),
   axis.text.x = element_blank(), 
   axis.text.y = element_blank(),
   axis.ticks = element_blank()
     )

gridExtra::grid.arrange(xdensity, blankPlot, scatterplot, ydensity,ncol=2, nrow=2, widths=c(4, 1.4), heights=c(1.4, 4))

gridExtra::grid.arrange(xdensity, blankPlot, scatter_gender_reg, ydensity,ncol=2, nrow=2, widths=c(4, 1.4), heights=c(1.4, 4))

```





