---
title: "FaceStroop Experiment Analysis"
author: "Mikkel Wallentin"
date: "16/02/2022"
output: html_document
---
#Get and clean FaceStroopExp data
```{r}
library(tidyverse)
library(lmerTest)

#Get FaceStroopData
datadir<-"./FaceStroopExp_data"


#Find files
files<-list.files(datadir,pattern='^FaceStroop_.+?csv',full.names=TRUE)

#Prepare an empty data frame for the data (also removes old version)
dataFS<-data.frame()

#How many datasets were there
n_datasets_raw<-length(files)
#Prepare a variable to monitor how many datasets we keep
n_datasets<-0
#Prepare a variable to monitor how many points we originally had
n_datapoints_raw<-0

#Loop to go through all files in the list
for(iii in 1:n_datasets_raw){
  
  #remove old loaded file to not risk importing it multiple times
  if(exists('data_temp')) rm(data_temp)
  
  #Load data
  data_temp<-read.csv(files[iii])
  
  #Take trials that have an RT (remove intertrial fixations)
  data_temp<- subset(data_temp,rt!='null' & trial_type=='image-keyboard-response')
  
  #Make a variable with the total number of loaded data points, to know how many have been discarded 
  n_datapoints_raw<-n_datapoints_raw+length(data_temp[,1])
  
  #Select data where at least 20 trials have been completed
  if(length(data_temp[,1])>20){
  
  #Some early data did not have a hunger measurement. Add a hunger variable for them and make them NAs
  if("hunger" %in% names(data_temp)){} else data_temp$hunger=NA
  
  #Some very ugly code because of script bug (you are able to select more than one value for hours since sleep and eat). 
  #We split the string, convert it from a list and turn the first value into an integer (a number)
  if(is.integer(data_temp$H_since_Eat)==FALSE){
    data_temp$H_since_Eat<-as.integer(unlist(strsplit(data_temp$H_since_Eat[1],","))[1])
    }
  if(is.integer(data_temp$H_since_Sleep)==FALSE){ 
    data_temp$H_since_Sleep<-as.integer(unlist(strsplit(data_temp$H_since_Sleep[1],","))[1])
    }
  
  #Make sure that data have the same number of columns
  if(length(colnames(data_temp))==33){
    #Only include dataset with more than 70 percent correct responses
    if(sum(data_temp$correct=='true')/length(data_temp$correct)>0.70){
      data_temp$correct_p<-mean(as.numeric(data_temp$correct=='true'))
      
      #Turn RT into numeric variable
      data_temp$rt<-as.numeric(data_temp$rt)
      data_temp$rt_median<-median(data_temp$rt)
      #Determine a cutoff for when response times are suspiciously long (e.g. 3 SDs above the mean)
      cutoff<-mean(data_temp$rt)+3*sd(data_temp$rt)
      #Filter data with too long RTs
      data_temp<- subset(data_temp,rt<cutoff)

      #If this is the first time around the loop, make the currently loaded data into the actual data
      if(length(colnames(dataFS))==0){
        dataFS=data_temp
        rm(data_temp)
        #counter to monitor included datasets
        n_datasets<-n_datasets+1
      }
      #Bind loaded data with actual data
      else {dataFS<-rbind(dataFS,data_temp)
      rm(data_temp)
        #counter to monitor included datasets
        n_datasets<-n_datasets+1
      }
    }
  }
  }
}


#A variable to monitor how many points we keep
n_datapoints<-length(dataFS[,1])

```


## Prepare variables for analysis
```{r}


#Make a variable that says if the condition was a Stroop condition or not (incongruency between word and face)
dataFS$stroop<-as.factor(dataFS$face!=dataFS$word)

#Limit the data to only correct answers. Response time for incorrect responses can be difficult to interpret.
#Note that this will make the data unsuitable for analyses on accuracy.
dataFS<- subset(dataFS,correct=='true')

#Turn id into factor - this can be used to monitor number of unique participants and to model random effects
dataFS$id <- dataFS$id %>% as.factor()

#Make a variable which has hour and minutes of the day as decimal variable
dataFS$hour2<-dataFS$hour+(dataFS$minute)/60
```

## Make a smaller dataframe for use when analysing mood etc, that were not repeated each session
```{r}
library(tidyverse)
#Summarise data in order to make smaller dataset with only one row per experimental session.
dataFSs<-dataFS %>% group_by(exp_id) %>% slice_head()
#Include a mean response time
temp<-dataFS  %>% group_by(exp_id) %>% summarise(rt_mean=mean(rt))
#Merge rt_mean with data
dataFSs<-merge(dataFSs,temp, by='exp_id')
rm(temp)
```


## It all starts with a circle and a triangle
```{r}
#Radius
r<-1
# Get x-values from minus 1 to plus 1
x<-seq(-1,1,0.00001)
# A circle with c(0,0) centre can be written with these two equations (following Pythagoras)
y1<-sqrt(r^2-x^2)
y2<--sqrt(r^2-x^2)
y<-c(y1,y2)
x<-c(x,x)
#Plotting the circle with sine and cosine values
pp=pi/4
plot(x,y,type='l')
lines(x=c(0,0),y=c(-1,1))
lines(x=c(0,cos(pp)),y=c(0,sin(pp)),col='darkgreen')
lines(x=c(cos(pp),cos(pp)),y=c(0,sin(pp)),col='darkblue')
text(x=c(-0.15+cos(pp)),y=c(0.5*sin(pp)),labels='sin(x)',col='darkblue')
lines(x=c(0,cos(pp)),y=c(sin(pp),sin(pp)),col='darkred')
text(x=c(0.5*cos(pp)),y=c(+0.15+sin(pp)),labels='cos(x)',col='darkred')
```

### Going beyond the circle, the sine and cosine functions can describe cycles in time

$$
y(t)=\beta{sin}(2\pi f t )
$$
$$
y(t)=\beta{cos}(2\pi f t )
$$
Beta is the amplitude (the height of the wave)
f is the frequency (the number of cycles per time unit)
t is time point



#### Plotting circadian sine and cosine waves
```{r}
#cycle frequency (in this case per hour) - one cycle per 24 hours
cf=1/24
#sample frequency per hour
fs=100
#Duration in hours
dur=24
#A time vector divided by fs
t = seq(0, dur, 1/fs)
#Make a sine wave (with amplitude =1) for each time point in t
u = sin(2*pi*cf*t)
#Make a cosine wave (with amplitude =1) for each time point in t
u2= cos(2*pi*cf*t)
#Plot the waves
plot(x=t,y=u, type='l',col='darkblue',xlab='time(hours)',ylab='Amplitude')
text(x=1+t[1000],y=-0.2+u[500],labels='y=sin(2*pi*cf*t)',col='darkblue')
lines(x=t,y=u2, type='l',col='darkred')
text(x=-1+t[500],y=-0.2+u2[500],labels='Y=cos(2*pi*cf*t)',col='darkred')
```
As can be seen, the sine function is a time shifted version of the cosine function and vice versa.
The time shift is also called phase. We can add a constant for the phase (phi) to the sine/cosine wave function.

$$
y(t)=\beta{sin}(2\pi f t + \phi)
$$

#### Plotting circadian sine wave with phase shift
```{r}
#cycle frequency (in this case per hour) - one cycle per 24 hours
cf=1/24
#sample frequency per hour
fs=100
#Duration in hours
dur=24
#A time vector divided by fs
t = seq(0, dur, 1/fs)
#a  phase shift of pi/2 radians (half a cycle)
phi=pi
#Make a sine wave (with amplitude =1)
u = sin(2*pi*cf*t)
#Make a sine wave (with amplitude =1), and phase shift
u2= sin(2*pi*cf*t+phi)
#Plot the waves
plot(x=t,y=u, type='l',col='darkblue',xlab='time(hours)',ylab='Amplitude')
text(x=1+t[1000],y=-0.2+u[500],labels='y=sin(2*pi*cf*t)',col='darkblue')
lines(x=t,y=u2, type='l',lty='dashed',col='darkblue')
text(x=-1+t[500],y=0.2+u2[500],labels='Y=sin(2*pi*cf*t+phi)',col='darkblue')
```
Using the trigonometric identity

$$
{sin(A+B)}={sin(A)cos(B)}-{cos(A)sin(B)}
$$
we can rewrite the sine function (including phase) as

$$
y(t)={\beta_1{sin}(2\pi f t)}+{\beta_2{cos}(2\pi f t)}
$$
where 
$$
{\beta_1}=\beta{cos(\phi)}, {\beta_2}=-\beta{sin(\phi)}
$$
We can use the rewritten sine function in a linear regression analysis, where we estimate the best fitting B1 and B2. This will yield a composite estimate of the amplitude and the phase of the data.

This will allow us to use sine and cosine waves to model a circadian rhythm, even if we don't know when it peaks. The amplitude will be given by

$$
{\beta}=\sqrt{\beta_1^2+\beta_2^2}
$$
The phase (phi) will be given by
$$
{\phi}=atan2(\beta_1,\beta_2)
$$

#### Use the sine and cosine waves to make predictors for the FaceStroop data

```{r}

#Make sine and cosine waves for each time point present in the data

dataFS$sinCirc<-sin(2*pi*cf*dataFS$hour2)
dataFS$cosCirc<-cos(2*pi*cf*dataFS$hour2)

# Plot the predictors for each data point in the data
plot(x=dataFS$hour2,y=dataFS$sinCirc,type='p',col='darkblue')
lines(x=dataFS$hour2,y=dataFS$cosCirc,type='p',col='darkred')
```





## Oscillation model

### Response time
```{r}
# Simple linear model with time of day predicting RT
modelFSrtHour<-lmer(rt~hour2+(1|id),data=dataFS)
# 2nd order polynomial model with time-squared predicting RT
modelFSrtHour2<-lmer(rt~poly(hour2,2)+(1|id),data=dataFS)
# Oscillation model with a sine and a cosine function predicting RT
modelFSrtCirc<-lmer(rt~sinCirc+cosCirc+(1|id),data=dataFS)

#Comparing the three different models to see which best explains the data
anova(modelFSrtHour,modelFSrtHour2,modelFSrtCirc)
```

#### Get fitted effects
```{r}
#Get info from model 1
m_temp<-summary(modelFSrtHour)
m_temp
#Make a variable with the fitted effect using the regression coefficients
dataFS$HourFit<-m_temp$coefficients[1,1]+m_temp$coefficients[2,1]*dataFS$hour2


#Get info from model 2
m_temp<-summary(modelFSrtCirc)
m_temp
#Make a variable with the fitted effect using the regression coefficients
dataFS$CircFit<-m_temp$coefficients[1,1]+m_temp$coefficients[2,1]*dataFS$sinCirc+m_temp$coefficients[3,1]*dataFS$cosCirc


#Get info from model 3
m_temp<-summary(modelFSrtHour2)
m_temp
#Make a variable with the fitted effect using the regression coefficients
dataFS$Hour2Fit<-m_temp$coefficients[1,1]+m_temp$coefficients[2,1]*dataFS$hour2+m_temp$coefficients[2,1]*dataFS$hour2^2


library(ggplot2)
#Make a fitted plot. Include a non-linear fit (geom_smooth) for comparison.
ggplot(dataFS,aes(x=hour2,y=rt))+geom_smooth()+geom_smooth(aes(x=hour2,y=HourFit),col='darkred')+geom_smooth(aes(x=hour2,y=CircFit),col='darkgreen')+labs(x='Time of day (hours)', y='Response time')
```

### Freshness
```{r}
# Freshness: Simple oscillation model
modelFSfreshCirc<-lmer(fresh~sinCirc+cosCirc+(1|id),data=dataFS)

#Get info from freshness model
m_temp<-summary(modelFSfreshCirc)
m_temp

#Make a variable with the fitted effect using the regression coefficients
dataFS$CircFreshFit<-m_temp$coefficients[1,1]+m_temp$coefficients[2,1]*dataFS$sinCirc+m_temp$coefficients[3,1]*dataFS$cosCirc



#Make a fitted plot. Include a non-linear fit (geom_smooth) for comparison.
ggplot(dataFS,aes(x=hour2,y=fresh))+geom_smooth()+geom_smooth(aes(x=hour2,y=CircFreshFit),col='darkgreen')+labs(x='Time of day (hours)', y='Tired/Fresh')
```
$$model=\beta_0+\beta_1 sin(x)+\beta_3 cos(x)$$

### Mood
```{r}
# Mood: Simple oscillation model
modelFSMoodCirc<-lmer(mood~sinCirc+cosCirc+(1|id),data=dataFS)

#Get info from mood model
m_temp<-summary(modelFSMoodCirc)
m_temp
#Make a variable with the fitted effect using the regression coefficients
dataFS$CircMoodFit<-m_temp$coefficients[1,1]+m_temp$coefficients[2,1]*dataFS$sinCirc+m_temp$coefficients[3,1]*dataFS$cosCirc

#Make a fitted plot. Include a non-linear fit (geom_smooth) for comparison.
ggplot(dataFS,aes(x=hour2,y=mood))+geom_smooth()+geom_smooth(aes(x=hour2,y=CircMoodFit),col='darkred')+labs(x='Time of day (hours)', y='Negative/Positive Mood')
```



### Compare time course of the three cycles
```{r}

#Make a fitted plot. Include a non-linear fit (geom_smooth) for comparison.
ggplot(dataFS,aes(x=hour2,y=fresh))+geom_smooth(aes(x=hour2,y=scale(CircFit)),col='darkblue')+geom_smooth(aes(x=hour2,y=scale(CircFreshFit)),col='darkgreen')+geom_smooth(aes(x=hour2,y=scale(CircMoodFit)),col='darkred')+labs(x='Time of day (hours)', y='RT/Fresh/Mood')
```






### Compare 1,2 and 3 cycle:
```{r}





```

















