---
title: "Portfolio 2"
author: "Studygroup 5 (Maja, Niels, Marton, Laurits & Sarah S.)"
date: "26/10/2021"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = T)

pacman::p_load(pastecs, tidyverse, readbulk, stringr, WRS2, car)

```

**variables**
Name: Subject identification (Factor)
Age: Age (Int)
Gender: Gender of the participant (Factor)
Condition: control = No surprising words, salient = there will be surpising words (Factor)
Word: The word being read (Character)
Reading_time: The reading time of that particular word. (Numeric) 

```{r}
df <- readbulk::read_bulk("logfiles", extension = ".csv")

df <- df %>% 
  rename(word_number = X)

view(df)
```

```{r}

mrc <- read_csv("MRC_database.csv")

log <- df %>% 
  mutate(word = str_to_upper(word))

df <- log %>% 
  inner_join(mrc)

df <- df %>% 
  mutate(
    var = if_else(is.na(lag(word)), 
                  TRUE, 
                  lag(word) != word)
    ) %>% 
  filter(var)

```

We need to examine whether or not our data is normally distributed in order to do t-tests on it. 
Therefore, we will do a Shapiro Wilk test on our data to get statistical evidence and also visualize it in a  histogram and a qq-plot.

```{r}
round(pastecs::stat.desc(cbind(df$reading_time), basic = FALSE, norm = TRUE), digits = 2)

df %>% 
  ggplot(aes(sample = reading_time)) +
    stat_qq() +
    stat_qq_line(colour = "red") +
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
    ggtitle("Reading_time, qq plot") +
    theme_bw()

df %>% 
  ggplot(aes(reading_time)) +
      geom_histogram(aes(y = ..density..), colour = "white", fill = "black") +
      stat_function(fun = dnorm, args = list(mean = mean(df$reading_time),
             sd = sd(df$reading_time)), colour = "red", size = 1) +
      ggtitle("Reading_time, histogram") +
      theme_bw()
  
```

From the shapiro Wilk test we can see that our data is not normally distributed, since the skew.2SE>1, kurt.2SE>1 and p<0.001. The qq-plot supports that our data is not normally distributed, because it does not follow the line; which is to be expected as we are looking at reaction times. Therefore, we can try to see whether or not transforming the data and removing outliers result in a normal distribution.

```{r}
filter_outlier <- df %>% 
  filter(reading_time > mean(reading_time)-sd(reading_time)*3) %>% 
  filter(reading_time < mean(reading_time)+sd(reading_time)*3) 

filter_outlier <- filter_outlier %>% 
  mutate(reading_time_log = log(reading_time), 
        reading_time_sqrt = sqrt(reading_time), 
        reading_time_divid = 1/reading_time
      )

```

Now we check if the transformed data is normally distributed:

```{r}

round(pastecs::stat.desc(cbind(filter_outlier$reading_time_log), basic = FALSE, norm = TRUE), digits = 2)

filter_outlier %>% 
  ggplot(aes(sample = reading_time_log)) +
    stat_qq() +
    stat_qq_line(colour = "red") +
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
    ggtitle("Reading_time, qq plot") +
    theme_bw()

filter_outlier %>% 
  ggplot(aes(reading_time_log)) +
      geom_histogram(aes(y = ..density..), colour = "white", fill = "black") +
      stat_function(fun = dnorm, args = list(mean = mean(filter_outlier$reading_time_log),
             sd = sd(filter_outlier$reading_time_log)), colour = "red", size = 1) +
      ggtitle("Reading_time_log, histogram") +
      theme_bw()

round(pastecs::stat.desc(cbind(filter_outlier$reading_time_sqrt), basic = FALSE, norm = TRUE), digits = 2)

filter_outlier %>% 
  ggplot(aes(sample = reading_time_sqrt)) +
    stat_qq() +
    stat_qq_line(colour = "red") +
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
    ggtitle("Reading_time_sqrt, qq plot") +
    theme_bw()

filter_outlier %>% 
  ggplot(aes(reading_time_sqrt)) +
      geom_histogram(aes(y = ..density..), colour = "white", fill = "black") +
      stat_function(fun = dnorm, args = list(mean = mean(filter_outlier$reading_time_sqrt),
             sd = sd(filter_outlier$reading_time_sqrt)), colour = "red", size = 1) +
      ggtitle("Reading_time_sqrt, histogram") +
      theme_bw()

round(pastecs::stat.desc(cbind(filter_outlier$reading_time_divid), basic = FALSE, norm = TRUE), digits = 2)

filter_outlier %>% 
  ggplot(aes(sample = reading_time_divid)) +
    stat_qq() +
    stat_qq_line(colour = "red") +
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
    ggtitle("reading_time_divid, qq plot") +
    theme_bw()

filter_outlier %>% 
  ggplot(aes(reading_time_divid)) +
      geom_histogram(aes(y = ..density..), colour = "white", fill = "black") +
      stat_function(fun = dnorm, args = list(mean = mean(filter_outlier$reading_time_divid),
             sd = sd(filter_outlier$reading_time_divid)), colour = "red", size = 1) +
      ggtitle("reading_time_divid, histogram") +
      theme_bw()

```
We can see that the transformed data is not normally distributed. 

Now we can explore if a relation exists between reading times and varies aspects of the words using correlation and scatter plots.

```{r}

round(pastecs::stat.desc(cbind(df$reading_time, df$nlet, df$kf_freq), basic = FALSE, norm = TRUE), digits = 2)

```

Now we plot the relations, in order to see if there is a relationship between the variables.  

Assumptions of parametric tests
1. Data are normally distributed
2. Variance is homogeneous across samples, groups, levels of a variable
3. Data are at least at the interval level
4. Data are independent from each other across participants or across sessions within participants

Since our data does not check the assumptions for a parametric test, we need to do a non-parametric correlation test. Thus, we use Spearman's correlation test.

```{r}
filter_outlier %>% 
  ggplot() +
  aes(reading_time, kf_freq) +
  geom_point() +
  geom_smooth(method = "lm")

cor.test(df$reading_time, df$kf_freq, method = "spearman")

filter_outlier %>% 
  ggplot() +
  aes(reading_time, nlet) +
  geom_point() +
  geom_smooth(method = "lm")

cor.test(df$reading_time, df$nlet, method = "spearman")

filter_outlier %>% 
  ggplot() +
  aes(reading_time, word_number) +
  geom_point() +
  geom_smooth(method = "lm")

cor.test(df$reading_time, df$word_number, method = "spearman")

```

From the scatterplot and the spearman's correlation test, we can see that there is no linear relation between the variables, except for the variable word_number. There is a statistical possibility that there is a linear relationship between word_number and reading_time, because p-value<0,05. 

Now we need to examine if our data meets the assumptions for a parametric t-test, but we need to check the assumptions separately for the two conditions, because they represent data from different groups. 
The assumptions: Assumes the dependent variable are normally distributed and that the variances are equal

We make a new dataframe containing the mean reading time for the word after the "salient word" and the "salient word", check for normality and a levene's test
```{r}

means_df <- df %>%
  filter(salience | lag(salience)) %>% 
  group_by(name) %>% 
  summarise(mean = mean(reading_time), condition = first(control)) %>% 
  mutate(condition = if_else(condition, "control", "salient"))

salient_means <- means_df %>% filter(condition != "control")
control_means <- means_df %>% filter(condition == "control")

```

Checking for normality
```{r}
round(pastecs::stat.desc(cbind(salient_means$mean), basic = FALSE, norm = TRUE), digits = 2)

salient_means %>% 
  ggplot(aes(sample = mean)) +
    stat_qq() +
    stat_qq_line(colour = "red") +
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
    ggtitle("salient_means, qq plot") +
    theme_bw()

salient_means %>% 
  ggplot(aes(mean)) +
      geom_histogram(aes(y = ..density..), colour = "white", fill = "black") +
      stat_function(fun = dnorm, args = list(mean = mean(salient_means$mean),
             sd = sd(salient_means$mean)), colour = "red", size = 1) +
      ggtitle("salient_means, histogram") +
      theme_bw()

round(pastecs::stat.desc(cbind(control_means$mean), basic = FALSE, norm = TRUE), digits = 2)

control_means %>% 
  ggplot(aes(sample = mean)) +
    stat_qq() +
    stat_qq_line(colour = "red") +
    labs(x = "Theoretical Quantiles", y = "Sample Quantiles") +
    ggtitle("control_means, qq plot") +
    theme_bw()

control_means %>% 
  ggplot(aes(mean)) +
      geom_histogram(aes(y = ..density..), colour = "white", fill = "black") +
      stat_function(fun = dnorm, args = list(mean = mean(control_means$mean),
             sd = sd(control_means$mean)), colour = "red", size = 1) +
      ggtitle("control_means, histogram") +
      theme_bw()
```
From the normality test, we can see that the data is normally distributed (skew.2SE<1, kurt.2SE<1, p-value>0,05). Thus, the data meets the first assumption for a parametric t-test. 

Levene's test
H0: no difference in the variance among the two groups
H1: a difference in the variance among the two groups
```{r}
my_data = stack(list(control_means=control_means$mean, salient_means=salient_means$mean))
leveneTest(values ~ ind, my_data)

```
Levene’s test is an inferential statistic used to evaluate the equality of variances for a variable determined for two or more groups. The p-value of the test is 0.1089, which is more than our significance level of 0.05. Therefore, we accept the null hypothesis and reject the alternative hypothesis and conclude that the variance among the two groups are equal.

Thus, our data meets the assumptions for a parametric t-test, and we can use the Welch two sample t-test. 
Our hypotheses:
H0 (null hypothesis) = No difference in the mean reading times in the two conditions of our reading experiment
H1 (alternative hypothesis) = There is a difference in the mean reading times in the two conditions of our reading experiment
```{r}

t.test(mean ~ condition, data = means_df)

```

On average, participants that were exposed to the salient text had no significantly larger reading time (M = 0.46), compared to participants who were exposed to the control text (M = 0.42), t(19) = -0.6, p>0.05. Therefore, we accept the null hypothesis and reject the alternative hypothesis, meaning that there is no significant difference in the mean reading times in the two conditions of our reading experiment. 
